Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2017-08-09 paper.tex
Kang-Kuen's edit

\begin{document}
\title{Motional Ground State Cooling Outside the Lamb-Dicke Regime}
\author{Yichao Yu}
\email{yichaoyu@g.harvard.edu}
\author{Nicholas R. Hutzler}
\altaffiliation{Present address: California Institute of Technology, Division of Physics, Mathematics, and Astronomy. Pasadena, CA, 91125}
\author{Jessie T. Zhang}
\author{Lee R. Liu}
\author{Kang-Kuen Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature, and
  large differential AC Stark shifts between the excited state and the ground state,
  we achieve a ground state preparation fidelity of $77(4)$\% after $100$ ms of cooling.
  Our technique includes addressing very high-order sidebands and
  fast modulation of the optical tweezer trap.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is applicable
  outside the Lamb-Dicke regime, for example in
  systems where tight confinement and low initial temperature are difficult to realize.
  This is particularly relevant for systems which are challenging to laser-cool,
  such as molecules and exotic atoms,
  and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematic of Raman sideband cooling.
    The Raman transitions have a one photon detuning $\Delta=25$ GHz from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atoms out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in the $|F=2,m_F=-2\rangle$ state
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematic of the cooling sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial direction are either addressing two neighboring cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards the end of the cooling sequence.
    All of the Raman pulses use Blackman pulse envelopes to reduce the off-resonant
    coupling to carrier and heating sidebands.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping state changing probability and Raman matrix element for axial direction.
    The range plotted covers $95$\% of the initial thermal distribution.
    (A) Axial motional state distribution after one optical pumping cycle
    for different initial $n$. Due to the large Lamb-Dicke parameter,
    there is a high probability of $n$ changing.
    (B) Matrix elements for Raman transition in the axial direction showing deviation from
    $\sqrt{n}$ scaling and multiple minima for different sideband orders.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in axial direction).
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_p1.pdf}
  \caption{(A) Radial Raman sideband spectrum of first order heating, first order cooling and
    second order cooling before and after Raman sideband cooling.
    (B,C) Rabi flopping on axis 3 (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue lines correspond to a ground state probability of $93$\% after cooling and
    the red lines correspond to a thermal distribution of $70$ $\mu$K before cooling.
    \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_p1.pdf}
  \caption{(A) Axial Raman sideband spectrum from first order heating to eighth order cooling
    before and after Raman sideband cooling.
    The data for the second and higher orders of cooling sidebands are taken with 150 $\mu$s
    pulse time and the rest are taken with 125 $\mu$s pulse time.
    (B,C) Rabi flopping on axial (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue and green lines corresponds to a ground state probability of $92$\% after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    The blue lines do not take into account the effect of decoherence due to resonance
    fluctuation. By comparing it to the green line in (C), which includes a $3$ kHz fluctuation,
    we can clearly see that this effect is the strongest for the post-cooling data on
    the axial heating sideband where the Rabi frequency is the lowest.
    The asymptotic values for the red lines are low also because of resonance fluctuation.
    \label{f-axial}}
\end{figure*}

Systems of trapped neutral atoms assembled from the bottom-up in optical tweezer arrays
are an exciting platform to study quantum information and quantum simulations~\cite{Schlosser2001,Weiss2004,Isenhower2010,Wilk2010,Kaufman2015,Labuhn2016,Murmann2015}.
The inherent single-particle detection and control combined with tunable interactions
are key to implementing neutral atom based quantum logic gates~\cite{Isenhower2010,Wilk2010},
novel quantum phases~\cite{Labuhn2016}, and single-photon switches~\cite{Dayan2008,Tiecke2014}.
Advances in real-time re-arrangement of optical tweezers enable rapid preparation of atoms
in large and complex geometries with high fidelity~\cite{Barredo2016,Endres2016}.
Adding quantum motional control of individual
atoms~\cite{Li2012,Kaufman2012,Thompson2013,Liu2017,Robens2017} enables
efficient coupling of single atoms to photonic crystal cavities~\cite{Thompson2013a},
high-fidelity single qubit gates~\cite{Wang2016},
and studies of the atomic Hong-Ou-Mandel effect~\cite{Kaufman2014}.

Extending optical tweezer array techniques to include polar molecules will open up a large range of
new applications that exploit their many long-lived internal states
and highly tunable interactions~\cite{DeMille2002,Ni2008,Gorshkov2011,Yan2013}.
Molecules could be assembled from atom pairs in the optical tweezer~\cite{Liu2017},
or directly loaded from magneto-optical traps
(MOTs)~\cite{Barry2014,Truppe2017SubDoppler,Anderegg2017}.
For either approach, preparing the constituent atoms or molecules in the lowest motional
quantum state is important for realizing long coherence times in quantum applications.

Preparation of single atoms in the motional ground state has been achieved in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
 using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems has been applied in the Lamb-Dicke (LD) regime, where
the size of the atomic wavefunction is much smaller than the wavelength of light used to address them.
Extending ground state cooling to systems outside the LD regime, such as light atoms or polar molecules, is challenging.
In this letter we demonstrate cooling of single sodium atoms in optical tweezers, initially outside of the LD regime, to the motional ground state.
We achieve a ground state probability of $P_0=77(4)$\% by utilizing cooling of
very high-order sidebands in a carefully optimized cooling sequence.
Our approach is general and opens up ground state cooling for other systems.

Our experiment begins by loading a single sodium atom into an optical tweezer from a MOT
for $250$ ms~\cite{Hutzler2017-LightShifts}, and has an overall repetition rate of $2.5$ Hz.
The tweezer is created by focusing a $700$ nm laser beam through an NA=$0.55$ objective to
create a focused beam with waist $0.7 \mu$m.
With $45$ mW of tweezer power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{430(4), 590(5), 69(1)\}\ \text{\text{kHz}}$
where axis z is along the weakly confined axial direction and axis x and y are along the
tightly confined radial directions.
After attempting to load an atom from the MOT, an image is taken with $1.5$ ms exposure time to determine if it was a success.
At the same time, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to an initial mean motional state along the three different axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.4(2),\, 2.5(2),\, 21(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
A beam with $\sigma^-$-polarization is used to initialize the atom into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping.
The hyperfine state detection is performed by pushing out atoms in the stretched state with a strong
beam on resonance with the $|F=2, m_F=-2\rangle$ to $|F'=3, m_{F'}=-3\rangle$ transition before
taking the second image to determine if the atom was pushed out.
The state preparation and detection have a combined fidelity of larger than $99$\%.

To further reduce the temperature of the single atom and
to achieve high ground state preparation fidelity, we apply Raman sideband cooling (RSC).
The relevant energy levels, cooling sequence, and Raman beam geometries for our system
are shown in Figure \ref{f-setup}. Briefly, the RSC consists of two steps:
driving a coherent Raman transition while removing motional quanta, followed by resetting the atom's internal state via optical pumping (OP).
RSC is applied repeatedly until the motional ground state is populated with high probability.

Specifically, Raman transitions are driven in Na between the hyperfine states
$|F=2, m_F=-2\rangle$ and $|F=1, m_F=-1\rangle$ in the presence of a $8.8$ G magnetic field to suppress vector light shifts~\cite{Kaufman2012,Thompson2013}.
Subsequently, an OP/repump pulse brings the atom back to $|F=2, m_F=-2\rangle$
via spontaneous emission.

It is important that the re-pumping step scatter as few photons as possible to minimize heating.
Therefore, we use a $\sigma^-$-polarized laser resonant with the D1 line
since $|F=2, m_F=-2\rangle$ is a dark state in this case~\cite{Grobner2017}\fxnote{cite ion}.
Compared to using a D2 repump beam, from which the $|F=2, m_F=-2\rangle$ state
could always scatter a photon via the excited $|F=3, m_F=-3\rangle$ state,
we find a reduction in the scattering rate of a factor of $130(20)$.

For an atom in motional level $n$, the OP could result in motional-state redistribution
as shown in Fig.~\ref{f-ld}A. The state redistribution probability
is approximately proportional to the effective LD parameter
$\eta^{\textrm{OP}}_{\textrm{eff}}=\sqrt{2n+1}\eta^{\textrm{OP}}$, where $\eta^{\textrm{OP}}=k z_0=k \sqrt{\hbar/2m\omega}$,
$k$ is the wavenumber of the optical pumping beam, $z_0$ is the zero-point wavefunction spread,
$m$ is the mass of the atom, and $\hbar$ is the reduced Planck constant.
Such heating presents a major challenge for efficient cooling of single Na atoms
which feature a light mass and short D-line wavelengths.
For our trap frequencies and beam geometry,
$\{\eta^{\textrm{OP}}_x,\eta^{\textrm{OP}}_y,\eta^{\textrm{OP}}_z\} = \{0.241(2), 0.206(1), 0.602(5)\}$.
Combined with a relatively large PGC temperature gives initial
$\{\eta^{\textrm{OP}}_{x,\textrm{eff}},\eta^{\textrm{OP}}_{y,\textrm{eff}},\eta^{\textrm{OP}}_{z,\textrm{eff}}\} = \{0.67(2), 0.50(1), 4.0(1)\}$.
For the weak axial direction, $(\eta^{\textrm{OP}}_{z,\textrm{eff}})^2>1$ is far outside of the LD regime.
As a result, the average change of motional state per OP step is large,
which creates a large uncertainty on the motional state at the end of a full RSC step, possibly leading to runaway heating.
For example, the average change in motional states over the initial distribution
is $2.4$ in the axial direction, which is significantly greater than $0.89-0.95$
in previous experiments of RSC of Rb and Cs~\cite{Li2012,Kaufman2012,Thompson2013,Liu2017}.

Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
The Raman transitions in our configuration have LD parameters of
$\{\eta^R_x,\eta^R_y,\eta^R_z\} = \{0.341(2), 0.291(1), 0.40(1)\}$,
which allow strong coupling to higher order sidebands for atoms in higher motional states
as calculated in Figure~\ref{f-ld}B.
The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$ similar to OP
where $\Delta k$ being the wave vector difference
between the two beams that drives the Raman transition.
To offset heating from OP initially, higher-order Raman cooling sidebands can be utilized
to remove more motional quanta in a single cooling pulse.
Since the coupling strengths of different orders do not reach minima for the same pulse duration,
using multiple orders of cooling sidebands avoids accumulations of population
near the coupling minima.

Taking the large motional-state changing heating and cooling sources into account,
it is not immediately clear that efficient cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search and
to find a robust cooling sequence\footnote{See supplemental material}.
In the simulation, we indeed observe a high heating rate due to the large LD parameters
and confirm that cooling with higher-order Raman sidebands can suppress heating.
In particular, we find that instead of cooling on only one sideband order repeatedly,
it is more efficient to alternate the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and 2nd- and 1st-orders for the radial directions
to minimize the accumulation of the atom in motional states that have zero Raman coupling.
The simulation also indicates that setting the coupling strength of each cooling sideband
to drive a Rabi $\pi$ pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in \ref{f-ld}B) can yield efficient cooling.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state, so the final cooling cycles utilize only the first-order sideband while alternating between the three different axes.

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed cooling sidebands (8th- and 7th-orders)
and decreasing the orders to the next pair after every $6$ to $15$ cycles.
This process is repeated until we reach the first order cooling sideband, and most of the
population is in the first few excited states. We then switch to cooling only on the
first order sideband with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
The radial cooling is performed similarly to the axial cooling with the initial cooling orders being
the 2nd and 1st order and switching to first order only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which can then be used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.

There are two additional important challenges in cooling single sodium atoms.
First, the high initial temperature populates high motional states,
meaning that the atoms sample the anharmonicity of the trap away from the center.
In order to address this, we drive the sidebands with a large Rabi frequency
and short pulses to Fourier broaden the spectrum coverage, and use Blackman pulse envelopes
to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}.
The anharmonicity in the radial directions is calculated to be 1kHz/n, meaning that a
Rabi frequency of at least $15$ kHz is needed to couple to $99$\% population up to $n=15$.
Second, the tweezer trap causes a large AC Stark shift (as large as $300$ MHz)
in the excited state. This creates a large, position-dependent OP detuning
and mixes the excited-state hyperfine levels, which reduces the optical pumping fidelity.
We address this by strobing the trapping light at 3 MHz during the whole cooling sequence,
similar to our loading and imaging process~\cite{Hutzler2017-LightShifts},
while leaving OP on with constant intensity.
Due to the large light shift, the OP is effectively off whenever the trap light switches back on.
Since the atom can only be addressed by the optical pumping light when the trap light is off,
the effect of light shifts on OP fidelity is suppressed.

>>>>

\begin{document}
\title{Motional Ground State Cooling Outside the Lamb-Dicke Regime}
\author{Y. Yu}
\email{yichaoyu@g.harvard.edu}
\author{N. R. Hutzler}
\altaffiliation{Present address: California Institute of Technology, Division of Physics, Mathematics, and Astronomy. Pasadena, CA, 91125}
\author{J. T. Zhang}
\author{L. R. Liu}
\author{T. Rosenband}
\author{K.-K. Ni}
\email{ni@chemistry.harvard.edu}
\affiliation{Department of Chemistry and Chemical Biology, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Department of Physics, Harvard University, Cambridge, Massachusetts, 02138, USA}
\affiliation{Harvard-MIT Center for Ultracold Atoms, Cambridge, Massachusetts, 02138, USA}

@@

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite  a  large Lamb-Dicke parameter, high initial temperature, and
  large differential light shifts between the excited state and the ground state,
  we achieve a ground state population of $81(4)$\% after $100$ ms of cooling. In addition, $10\,\%$ of initial atom population escapes the trap during cooling.
  Our technique includes addressing  high-order sidebands, , where several motional quanta are removed by a single laser pulse,  and
  fast modulation of the optical tweezer intensity.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is possible %outside the Lamb-Dicke regime, 
  where tight confinement and low initial temperature are difficult to realize.
%  This is particularly relevant for systems which are challenging to laser-cool, such as molecules and exotic atoms, and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

%Systems of trapped neutral atoms assembled from the bottom-up in an array of optical tweezers are a promising  platform to study quantum information and quantum simulations~
Trapped neutral atoms, assembled  in an array of optical tweezers, are a promising  platform to study quantum information and quantum simulations~
\cite{Schlosser2001,Weiss2004,Isenhower2010,Wilk2010,Kaufman2015,Labuhn2016,Murmann2015}.
The inherent single-particle detection and control, combined with tunable interactions,
allow implementations of neutral-atom-based quantum logic gates~\cite{Isenhower2010,Wilk2010},
novel quantum phases~\cite{Labuhn2016}, and single-photon switches~\cite{Dayan2008,Tiecke2014}.
Advances in real-time re-arrangement of optical tweezers enable rapid preparation of atoms
in large and complex geometries~\cite{Barredo2016,Endres2016}.
Quantum motional control of individual
atoms~\cite{Li2012,Kaufman2012,Thompson2013,Liu2017,Robens2017} enable
studies of the atomic Hong-Ou-Mandel effect~\cite{Kaufman2014},
high-fidelity single qubit gates~\cite{Wang2016},
and efficient coupling of single atoms to photonic crystal cavities~\cite{Thompson2013a}.

Extending optical tweezer arrays  to include polar molecules will open up a  range of
new applications that exploit  long-lived molecular internal states
and  tunable long range interactions~\cite{DeMille2002,Gorshkov2011,Yan2013}.
Molecules could be assembled from atom pairs in  optical tweezers~\cite{Liu2017},
or directly loaded from magneto-optical traps
(MOTs)~\cite{Barry2014,Truppe2017SubDoppler,Anderegg2017}.
For either approach, preparing the constituent atoms or molecules in the lowest motional
quantum state is important to realize long coherence times in quantum applications.

Preparation of single atoms in the motional ground state has been achieved in optical tweezers~\cite{Kaufman2012,Thompson2013,Liu2017,Robens2017}
 using Raman sideband cooling (RSC)~\cite{Monroe1995,Kerman2000,Han2000}.
However, RSC in these systems take place in the Lamb-Dicke (LD) regime, where
the spread of the initial wavefunctions is much smaller than the wavelength of light used to address them.
Systems with a large wavefunction, either due to small mass or high initial temperature,
fall outside of the LD regime and undergo strong recoil heating during RSC.
Extending RSC beyond the LD regime opens  the possibility
for ground state cooling of systems such as light atoms and %systems with poor cooling such as 
molecules that can be laser-cooled.
In this letter we demonstrate cooling of single sodium atoms in optical tweezers,
initially outside of the LD regime, to the motional ground state.
We achieve a 3D ground state population of $P_0=81(4)$\% by  cooling via
 high-order sidebands in a carefully optimized cooling sequence.
Our approach is general and applicable to %opens up ground state cooling for 
other systems.

Our experiment has an overall repetition rate of $2.5$ Hz and
begins by loading a single sodium atom into optical tweezers from a MOT~\cite{Hutzler2017-LightShifts}.
%for $250$ ms~\cite{Hutzler2017-LightShifts}.
The tweezers are created by focussing a $700$ nm wavelength laser beam through an NA=$0.55$ objective to an elliptical waist with radii $\{w_{0,x},w_{0,y}\}\approx \{0.79,0.57\}\,\mu$m.
For $45$ mW of optical power, the trapping frequencies along the three axes are
$\{\omega_x,\omega_y,\omega_z\}/2\pi = \{430(4), 590(5), 69(1)\}\ \text{\text{kHz}}$,
where  $z$ is the weakly confined axial direction,
and  $x$ and $y$ are the tightly confined radial directions.
After each single atom loading attempt, an image is taken with a $1.5$ ms exposure time
to determine success.
During  imaging, the atom is cooled via polarization gradient cooling (PGC),
which reduces the temperature of the single atom to $70$ $\mu$K,
corresponding to mean motional states along the three trap axes of
$\{\bar n_x, \bar n_y, \bar n_z\}=\{3.4(2),\, 2.5(2),\, 21(2)\}$, and therefore initial 3D ground state probability of $P_0=0.3$\%.
The atom is then initialized  into
the $|F=2, m_F=-2\rangle$ stretched state via optical pumping (OP).
Each experimental cycle ends with a second imaging sequence that measures the hyperfine state of the atom as $|F=2\rangle$ or $|F=1\rangle$~\footnote{The final hyperfine state sensitive detection employs a strong beam that is resonant with the $F=2$ to $F'=3$ cycling transition, to heat the $F=2$ population out of the trap.  Therefore, any atoms which survive this heat out procedure are interpreted as having been in the $F=1$ state.}.%a repetition of the experiment that was initially bright and remained bright after the pulse sequence is interpreted as a final F=1 atomic state.}.

%\footnote{We were able to load into larger tweezers with waist $> 1$ micron with an unmodulated tweezer and MOT beams, though the loading was only a few percent efficient.}
%, while a bright image that turned dark is interpreted as F=2.   
%The hyperfine state detection is performed by pushing out atoms in the stretched state with a strong beam on resonance with the $|F=2, m_F=-2\rangle$ to $|F'=3, m_{F'}=-3\rangle$ transition before taking the second image to determine if the atom was pushed out. The state preparation and detection have a combined fidelity of larger than $99$\%.

To  reduce the temperature of the single atom and
 achieve high ground state population, we apply Raman sideband cooling~\cite{Monroe1995, Kaufman2012}.
The  energy levels, cooling sequence, and Raman beam geometries 
are shown in Fig.~\ref{f-setup}. RSC consists of two steps:
driving a coherent Raman transition  to remove motional quanta, followed by resetting the atom's internal state via OP.
These  steps are  repeated, until the motional ground state is populated with high probability.

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{Single Na atom Raman sideband cooling scheme and sequence. (A)
    The Raman transitions between $|2,-2;n\rangle$ and $|1,-1;n+\Delta n\rangle$ have a one-photon detuning $\Delta=25$ GHz below the $3^2S_{1/2}$ to $3^2P_{3/2}$ transition. Two-photon detuning, $\delta$, is defined relative to the $\Delta n=0$ carrier transition. For optical pumping, we use two $\sigma^-$ polarized transitions, one   to  pump the atom state out of $|1,-1\rangle$ via $3^2P_{3/2}$ and one  to pump atoms out of $|2,-1\rangle$ via $3^2P_{1/2}$
     to minimize heating on the atom in the $|2,-2\rangle$ state.
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezers and bias magnetic field.  Raman beams 1 and 4 address the radial $x$ mode. Raman beams 1 and 2 address the radial $y$ mode. Raman beams 3 and 4 address the axial $z$ mode, which also couples to the radial $y$ mode.
    (C) Schematic of the cooling pulse sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ sideband pulses.
    The four axial pulses address two sideband orders.
    The two pulses in each radial direction  either address $\Delta n=-2$ and $\Delta n=-1$
    or have different durations to drive $\Delta n=-1$, when most of the population is below $n=3$
    at the end of the cooling sequence.
    The Raman cooling and spectroscopy pulses have Blackman  envelopes to reduce  off-resonant
    coupling, while the measurement Rabi pulses in Fig.~\ref{f-radial}(B,C) and \ref{f-axial}(B,C) have square envelopes to simplify analysis.
    \label{f-setup}}
\end{figure*}

Specifically, Raman transitions are driven in Na between the hyperfine states
$|2, -2\rangle$ and $|1, -1\rangle$ in the presence of an $8.8$ G magnetic field. The external field is
orthogonal to the effective magnetic field of the tweezer, to reduce  vector light shifts~\cite{Kaufman2012,Thompson2013}.
Subsequently, an OP pulse, which consists of two frequencies  both with $\sigma^-$-polarization, is applied  to bring the atom out of the $F=1$ manifold and back to  $|2, -2\rangle$ via spontaneous emission (Fig.~\ref{f-setup}A).
%brings the atom back to $|F=2, m_F=-2\rangle$ via spontaneous emission.
%, which consists of two frequencies with $\sigma^-$-polarization for the  to bring the atom out of $F=1$ manifold and back to  $|F=2, m_F=-2\rangle$ via spontaneous emission.
It is important that the optical pumping step scatter as few photons as possible to minimize recoil heating. %The OP consists of driving a $|F=1, m_F=-1\rangle$ into the $F=2$ manifold via the transition of $3P_{3/2} D2 line$
Therefore, to optically pump atoms from $|2,-1\rangle$ into $|2, -2\rangle$, we apply light resonant with  $3^2P_{1/2}$~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3^2P_{3/2}$, from which the $|2, -2>$ state could always scatter a photon via the excited $|F'=3, m_F'=-3>$ state.} because $|2, -2\rangle$ is a dark state to $\sigma^-$-polarized light~\cite{Monroe1995, Grobner2017}.
%~\footnote{We find a reduction in the scattering rate by a factor of $130(20)$, as compared to using an OP resonant with $3P_{3/2}$, from which the $|2, -2 \rangle$}

\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping motional-state redistribution and  Raman coupling for large  Lamb-Dicke parameters for the axial direction ($z$). The range plotted covers $95$\% of the initial thermal distribution.
   % The range plotted covers $95$\% of the initial thermal distribution.
    (A) Motional state distribution after one optical pumping cycle
    for different initial n, $n_{\textrm{init}}$. Due to photon-recoil and the large Lamb-Dicke parameter, $\eta^{OP}_z=0.6$,
    there is a high  $n$ changing probability.         (B) Matrix elements for Raman transition deviate from
    $\sqrt{n}$ scaling with multiple minima. During cooling, we utilize the fact that higher motional states couple most effectively to sidebands with large $|\Delta n|$.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in the axial direction).
    \label{f-ld}}
\end{figure}

For an atom starting in motional level $n_{\textrm{init}}$, the probability of recoil during the OP sequence results in
a motional-state redistribution  is shown in Fig.~\ref{f-ld}A.
The  probability is approximately proportional to the square of effective LD parameter
$\eta^{\textrm{OP}}_{\textrm{eff}}=\sqrt{2n+1}\eta^{\textrm{OP}}$~\cite{Kaufman2012},
where the optical pumping LD parameter $\eta^{\textrm{OP}} = z_0\frac{2\pi}{\lambda}$ is the ratio of the zero-point wavefunction spread $z_0 = \sqrt{\hbar/2m\omega}$ to the reduced wavelength $\lambda$.
Here, $m$ is the mass of the atom, $\omega$ is the trap frequency,
and $\hbar$ is the reduced Planck constant.
Such heating presents a major challenge for efficient cooling of single Na atoms
which feature a low mass (23 amu) and short wavelengths.
For our configuration, 
$\{\eta^{\textrm{OP}}_x,\eta^{\textrm{OP}}_y,\eta^{\textrm{OP}}_z\} = \{0.241(2), 0.206(1), 0.602(5)\}$.
A relatively large PGC temperature results in initial $\{\eta^{\textrm{OP}}_{x,\textrm{eff}},\eta^{\textrm{OP}}_{y,\textrm{eff}},\eta^{\textrm{OP}}_{z,\textrm{eff}}\} = \{0.67(2), 0.50(1), 4.0(1)\}$.
The weak axial direction %, $(\eta^{\textrm{OP}}_{z,\textrm{eff}})^2>1$ 
is far outside of the LD regime ($\eta^2\ll 1$), and the probability of recoil heating per OP step is large. %~\footnote{For example, the average change in the motional states for the initial distribution is $2.4$ in the axial direction, which is significantly greater than $0.89-0.95$in previous experiments of RSC of Rb and Cs~\cite{Li2012,Kaufman2012,Thompson2013,Liu2017}.},
This creates a large uncertainty on the motional state at the end of a full RSC step, possibly leading to runaway heating.

Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$,
where the wave vector is replaced by the wave vector difference $\Delta k$
between the two beams that drive the Raman transition.
The Raman transitions in our configuration have LD parameters of
$\{\eta^R_x,\eta^R_y,\eta^R_z\} = \{0.341(2), 0.291(1), 0.40(1)\}$.
To offset heating from OP initially, higher-order Raman  sidebands ($\Delta n <-1$)  can be utilized
to remove several motional quanta in a single cooling pulse (see Fig.~\ref{f-ld}B).
Because the coupling strengths of different orders do not reach minima for the same state of motion $n$,
using multiple orders of sidebands avoids accumulations of population
near the coupling minima.

Taking the large motional-state changing heating and cooling sources
 into account,
it is not immediately clear that efficient cooling can be achieved.
We therefore use a Monte-Carlo simulation to guide our search~\cite{Dalibard1992}. 
%and to find a robust cooling sequence~\footnote{See supplemental material}.
In particular, we find that alternating the cooling pulses (Fig.~\ref{f-setup}C) between two
neighboring orders for the axial direction and $\Delta n=-2$ and $\Delta n=-1$ for the radial directions
eliminates the accumulation of population in motional states that have zero Raman coupling, which would halt the cooling process.
The simulation also indicates that setting the coupling strength of each  sideband
to drive a Rabi $\pi$-pulse corresponding to the maximum matrix element motional state
(i.e. the maxima in Fig.~\ref{f-ld}B)  yields efficient cooling, initially.
The efficiency of cooling on higher-order sidebands diminishes
as the atom approaches the ground state, so the final  cycles utilize only the $\Delta n=-1$ sideband while alternating between the three  axes.

Guided by the simulation results,
we construct our axial cooling sequence by starting at the two highest
observed  sideband orders ($\Delta n=-8$ and $\Delta n=-7$)
and decreasing the orders to the next pair after every $6$ to $15$ cycles.
This process is repeated until we reach $\Delta n=-1$, and most of the
population is in the first few excited states. We then switch to cooling only on $\Delta n=-1$ with two different pulse lengths in order to efficiently cool atoms in the
few remaining motional states.
Radial cooling is performed similarly to the axial cooling with the initial cooling of $\Delta n=-2$ and $\Delta n=-1$ and switching to $\Delta n=-1$ only after $20$ to $30$ pulses.
This sequence gives good initial cooling performance, which can then be used to measure experimental
parameters, including the Rabi rates of the Raman beams and optical pumping rates,
in order to optimize the sequence further.

There are two additional important challenges in cooling single sodium atoms.
First, the high initial temperature populates high motional states,
causing the atoms to sample the anharmonicity of the trap away from the center. Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.  We find $A_{i,n}=\frac{3n \hbar}{4 \pi m d_i^2}$  where $d_i=w_{0,x},w_{0,y}$ for the radial directions and $d_z=\pi w_{0,x}w_{0,y}/\lambda$.  Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{1.1, 2.0, 0.16\}n$~kHz. For mixed states, this  broadens high-order sidebands due to the $n$-dependence of the transitions. 
In order to mitigate this, we drive the sidebands with a large  Rabi frequency
and short pulses to Fourier broaden the spectrum. %, and use Blackman pulse envelopes to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}. 
However, anharmonicity  may still shift high-$n$ states out of resonance and prevent effective cooling.
%Anharmonicity may be defined as $A_{i,n}=(E_{i,n+1}-E_{i,n})/h - \omega_i/(2\pi)$ for each trap axis $i$, and calculated from the quartic term of the optical tweezers via perturbation theory.  We find $A_{i,n}=\frac{3n \hbar}{2 m d_i^2}$  where $d_i=w_0$ for the radial directions and $d_z=\pi w_0^2/\lambda$  Numerically, $\{A_{x,n},A_{y,n},A_{z,n}\}=\{1.3, 1.2, 0.2\}n$~kHz.  Anharmonicity broadens the high-order axial transitions and may shift high-$n$ states out of resonance to prevent effective cooling.

%The anharmonicity in the radial directions is calculated to be 1kHz/n, meaning that a Rabi frequency of at least $15$ kHz is needed to couple to $99$\% population up to $n=15$.
